\def\Nst#1{$#1^{st}$}\def\Nnd#1{$#1^{nd}$}\def\Nrd#1{$#1^{rd}$}\def\Nth#1{$#1^%
{th}$}
\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\newcommand{\enquote}[1]{``#1''}

\bibitem{Angluin:regset}
Angluin, D., \emph{Learning regaular sets from queries and counterexamples},
  Information and Computation \textbf{75} (1987), pp.~87--106.

\bibitem{Angluin01}
Angluin, D., \emph{Queries revisited}, in: N.~Abe, R.~Khardon and T.~Zeugmann,
  editors, \emph{Algorithmic Learning Theory, 12th International Conference
  2001},  Lecture Notes in Computer Science  \textbf{2225} (2001).

\bibitem{Cetal:bandera}
Corbett, J., M.~Dwyer, J.~Hatcliff, S.~Laubach, C.~Pasareanu, Robby and
  H.~Zheng, \emph{Bandera : Extracting finite-state models from java source
  code}, in: \emph{Proc.\ 22nd Int.\ Conf.\ on Software Engineering}, 2000.

\bibitem{FJJV:scp97}
Fernandez, J.-C., C.~Jard, T.~J{\'e}ron and C.~Viho, \emph{An experiment in
  automatic generation of test suites for protocols with verification
  technology}, Science of Computer Programming \textbf{29} (1997).

\bibitem{SPIN}
Gr\'egoire, J.-C., G.~J. Holzmann and D.~A. Peled, editors, \enquote{The Spin
  Verification System,}  DIMACS series  \textbf{32}, American Mathematical
  Society, 1997, iSBN 0-8218-0680-7, 203p.

\bibitem{GPY:adaptive}
Groce, A., D.~Peled and M.~Yannakakis, \emph{Adaptive model checking}, in:
  J.-P. Katoen and P.~Stevens, editors, \emph{Proc.\ {TACAS '02}, \Nth{8} Int.\
  Conf.\ on Tools and Algorithms for the Construction and Analysis of Systems},
   Lecture Notes in Computer Science  \textbf{2280} (2002), pp. 357--370.

\bibitem{HHNS:modelgeneration}
Hagerer, A., H.~Hungar, O.~Niese and B.~Steffen, \emph{Model generation by
  moderated regular extrapolation}, in: R.-D. Kutsche and H.~Weber, editors,
  \emph{Proc.\ {FASE '02}, \Nth{5} Int.\ Conf.\ on Fundamental Approaches to
  Software Engineering},  Lecture Notes in Computer Science  \textbf{2306}
  (2002), pp. 80--95.

\bibitem{Harel:statemate}
Harel, D., H.~Lachover, A.~Naamad, A.~Pnueli, M.~Politi, R.~Sherman,
  A.~Shtull-Trauring and M.~Trakhtenbrot, \emph{{STATEMATE}: A working
  environment for the development of complex reactive systems}, IEEE Trans.\ on
  Software Engineering \textbf{16} (1990), pp.~403--414.

\bibitem{Holzmann:spin00}
Holzmann, G., \emph{Logic verification of {ANSI-C} code with {SPIN}}, in:
  K.~Havelund, J.~Penix and W.~Visser, editors, \emph{{SPIN} Model Checking and
  Software Verification: Proc. \Nth{7} Int.\ {SPIN} Workshop},  Lecture Notes
  in Computer Science  \textbf{1885} (2000), pp. 131--147.

\bibitem{HungarNS03}
Hungar, H., O.~Niese and B.~Steffen, \emph{Domain-specific optimization in
  automata learning}, in: \emph{Proc.\ \Nth{15} Int.\ Conf.\ on Computer Aided
  Verification}, 2003, to appear.

\bibitem{Uppaal:nutshell}
Larsen, K., P.~Pettersson and W.~Yi, \emph{{UPPAAL} in a nutshell}, Software
  Tools for Technology Transfer \textbf{1} (1997).

\bibitem{SEG:testtools}
Schmitt, M., M.~Ebner and J.~Grabowski, \emph{Test generation with autolink and
  testcomposer}, in: \emph{Proc.\ 2nd Workshop of the SDL Forum Society on SDL
  and MSC - SAM'2000}, 2000.

\end{thebibliography}
