\documentclass{entcs}
\usepackage{entcsmacro}
\begin{document}
\begin{frontmatter}
  \title{\bf Preface}
%%\maketitle
\end{frontmatter}

This volume contains the proceedings of the Third international Workshop
on Software Verification and Validation (SVV). SVV took place  on
October 31, 2005 as a pre-conference satellite workshop of the
Seventh International Conference on Formal Engineering Methods (ICFEM) that
took place at the University of Manchester (UK), November 1---4, 2005.

The goal of SVV is to promote discussion on novel methodologies for
software verification as well as study novel combinations of (known)
methodologies. An outstanding example of a combined verification
methodology is the recent research direction that combines abstraction
(of infinite-state programs into finite-state ones) with model checking
(of finite-state systems). There is a growing conviction in the research
community that such hybrid methodologies are imperative for the process
of analyzing full-fledged software systems. 

The SVV 2005 program consisted of 6 exciting papers.  The topics
addressed by the accepted contributions covered theoretical techniques
as well as practical methods/case studies for verification of software
systems. Prof. Farn Wang from the National Taiwan University
delivered an invited talk titled ``Automatic Test Case Generation with
Region-Related Coverage Annotations for Real-time Systems''.

We would like to thank the program committee and all the referees for
their assistance in selecting the papers and the local organizers for
providing tremendous help with the coordination of the workshop.

\vspace{15pt}
\noindent ---\\
{\em Supratik Mukhopadhyay, Abhik Roychoudhury and Zijiang Yang}.\\
9 Jan 2006.
\end{document}


