\documentclass{entcs} 
\usepackage{prentcsmacro}
\usepackage{graphicx}
\usepackage{xspace}

\input{macros}

\newcommand{\Nat}{{\mathbb N}}
\newcommand{\Real}{{\mathbb R}}
\def\lastname{Castelluccia Mongiello Ruta Totaro}
\begin{document}
\begin{frontmatter}
  \title{WAVer: A Model Checking-based Tool to Verify Web Application Design}  
  \author{D. Castelluccia\thanksref{castellucciamail}}
  \author{M. Mongiello\thanksref{mongiellomail}}
  \author{M. Ruta\thanksref{rutamail}}
  \author{R. Totaro\thanksref{totaromail}} 
  \address{Dipartimento di Elettrotecnica ed Elettronica\\Politecnico di Bari\\ I-70125 Bari, Italy}
    \thanks[castellucciamail]{Email:
    \href{mailto:d.castelluccia@virgilio.it} {\texttt{\normalshape
        d.castelluccia@virgilio.it}}}
		\thanks[mongiellomail]{Email:
    \href{mailto:mongiello@poliba.it} {\texttt{\normalshape
        mongiello@poliba.it}}}
    \thanks[rutamail]{Email:
    \href{mailto:m.ruta@poliba.it} {\texttt{\normalshape
        m.ruta@poliba.it}}}
    \thanks[totaromail]{Email:
    \href{mailto:r.totaro@poliba.it} {\texttt{\normalshape
        r.totaro@poliba.it}}}


%\pagestyle{headings}
%In order to omit page numbers and running heads
%please change this line to
%\pagestyle{empty}
%and change the first command line too, see above.


\begin{abstract}
Web Applications are becoming more and more widespread and efficient, then an increase of their reliability is now strongly required. Hence methods to support design and automatically perform validation of a Web Application (WA) could be helpful. In this paper we present WAVer, a prototype tool for performing the verification of a WA design by means of Symbolic Model Checking techniques. The tool first performs the modeling of the WA and furthermore verify it by means of a model checker. Specifically, the mathematical model of the WA is represented by a Finite State Machine (FSM). Then, by using the \ctl formal language, we formalize basic criteria to establish correctness of the application. The prototype system we have implemented embeds a component which automatically imports WA design from a UML tool; \ctl specifications are added and translated as source code for NuSMV model checker. Finally, the checker performs verification: if there is a violation of specifications, NuSMV allows to locate errors in WA design and appropriate adjustments are carried out.
\end{abstract}

\begin{keyword}
  web application, verification, model checking.
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{introduction}
\input{introduction}

\section{Model Checking}\label{modelchecking}
\input{modelchecking}

\section{Proposed model}\label{model}
\input{model}

\section{From XMI to SMV}\label{xmi}
\input{xmi}

\section{Related Work}\label{related}
\input{related}

\section{Conclusion}
\input{conclusion}

\bibliographystyle{plain}
\bibliography{long,model}

\end{document}