\documentclass{entcs}
\usepackage{entcsmacro}
\begin{document}
\begin{frontmatter}
  \title{Preface}
%%\maketitle
\end{frontmatter}

This volume contains the proceedings of the First international Workshop
on Software Verification and Validation (SVV). SVV took place  on
December 14, 2003 as a post-conference satellite workshop of the
Nineteenth International Conference on Logic Programming (ICLP) that
took place at the Tata Institute of Fundamental Research (TIFR), Mumbai,
India, December 9--13, 2003.

The goal of SVV is to promote discussion on novel methodologies for
software verification as well as study novel combinations of (known)
methodologies. An outstanding example of a combined verification
methodology is the recent research direction that combines abstraction
(of infinite-state programs into finite-state ones) with model checking
(of finite-state systems). There is a growing conviction in the research
community that such hybrid methodologies are imperative for the process
of analyzing full-fledged software systems.

SVV 2003 received 14 submissions out of which 9 papers were selected for
presentation at the workshop. The papers were subjected to rigorous
review by at least three independent reviewers. In addition to the
standard criteria for acceptability, contributions have also been
selected on the basis of their conceptual significance in neighboring
areas. The topics addressed by the accepted contributions ranged from
theoretical techniques and practical methods to case studies for
verification of conventional and embedded software systems.
Dr Ramesh Bharadwaj from the Naval Research Laboratories delivered an
invited talk titled `` Rigorous Methods for Software Construction:
Retrofitting "Engineering" into Software Engineering''.

We would like to thank the program committee and all the referees for
their assistance in selecting the papers and the local organizers for
mastering the coordination of the workshop as well as the ICLP.

\vspace{5pt}
Sandro Etalle, Supratik Mukhopadhyay and Abhik Roychoudhury
\end{document}


